Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-11-15 ms.tex
brian revisions

\label{sec:sample}
%AG:It is unclear whether if/how the Nuker profile will show up in
%this parametrized study.
Our sample includes 61 early-type galaxies from
\citetalias{WangMerritt:2004a}, as represent a subset of a larger
sample of \citet{FaberTremaine+:1997a} with Hubble WFPC2 imaging.  We
focus on elliptical galaxies because we are interested in isolating
SMBHs fed by the steady mass loss from the relatively old stellar
population, instead of fresh gas flowing into the sphere of influence
from larger scales.

@@

The density of gas at the stagnation radius $\rho(r_{\rm s})$ can be estimated by using equations (\ref{eq:stag_simple}) and (\ref{eq:mdot_analytic}) in conjunction with an alternative estimate of the mass accretion rate as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}(\rs) = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$, viz.
\begin{align}
  &\dot{M}\simeq\frac{(4 \pi/3) \rs^3 \rho(\rs)}{t_{\ff}(\rs)},
  \label{eq:mdot_gas}
\end{align}
from which one derives 
\begin{align}
  \rho(\rs)\simeq
  \begin{cases}
    3 \times 10^{-24} \Mbheight^{-0.14} v_{500}^{-1}  \eta \,
    \, {\rm g \, cm^{-3}}& \text{core}\\
    6 \times 10^{-24}  \Mbheight^{-0.56} v_{500}    \eta \,\, {\rm g \,cm^{-3}} & \text{cusp}
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

@@

However, \citealt{McCourt+12} argue that even if cooling is exactly balanced by heating everywhere, there is still a local thermal instability that operates to turn hot gas into cold clouds.  However, this local instability only causes a significant local drop-out of mass from the hot medium (to low-T clouds), if the cooling timescale,
\be
t_{\rm cool} \equiv \frac{3n kT/2}{\dot{q}_{\rm cool}}
\ee 
is locally shorter than the approximately 10 times the free-fall timescale $t_{\rm ff}$.

@@

>>>>

\label{sec:sample}
%AG:It is unclear whether if/how the Nuker profile will show up in
%this parametrized study.
{\bf BDM: we may show plot of NUKER masses, $\Gamma$'s, etc here to (1) motivate the parametric study; (2) in individual cases we may want to compare solutions to specific galaxies.}

Our sample includes 61 early-type galaxies from
\citetalias{WangMerritt:2004a}, as represent a subset of a larger
sample of \citet{FaberTremaine+:1997a} with Hubble WFPC2 imaging.  We
focus on elliptical galaxies because we are interested in isolating
SMBHs fed by the steady mass loss from the relatively old stellar
population, instead of fresh gas flowing into the sphere of influence
from larger scales.

@@

The stellar density at the stagnation radius (eq.~[\ref{eq:stag_simple}]) is usefully approximated as
\begin{eqnarray}
  && \rho_{\star}(r_{\rm s}) \approx \nonumber \\
&& \begin{cases}
    1.0\times 10^{-18}M_{\bullet,8}^{-1.2}v_{500}^{2}\chi^{2}\,{\rm g\,cm^{-3}}
& \text{core} \\
    1.1 \times 10^{-18} M_{\bullet,8}^{-1.6}v_{500}^{4}\chi^{4}
      \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}

The density of gas at the stagnation radius $\rho(r_{\rm s})$ can be roughly estimated by using equations (\ref{eq:stag_simple}) and (\ref{eq:mdot_analytic}) in conjunction with an alternative estimate of the mass accretion rate as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}(\rs) = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$, viz.
\begin{align}
  &\dot{M}\simeq\frac{(4 \pi/3) \rs^3 \rho(\rs)}{t_{\ff}(\rs)},
  \label{eq:mdot_gas}
\end{align}
from which one derives 
\begin{align}
  \rho(\rs)\simeq
  \begin{cases}
    3 \times 10^{-24} \Mbheight^{-0.14} v_{500}^{-1}  \eta \,
    \, {\rm g \, cm^{-3}}& \text{core}\\
    6 \times 10^{-24}  \Mbheight^{-0.56} v_{500}    \eta \,\, {\rm g \,cm^{-3}} & \text{cusp}
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

Radiative cooling can have an important on even the qualitative features of the flow, because initially hydrostatic gas that cools at constant pressure $P \propto nT$ according to $\dot{q}_{\rm cool} \propto T^{-2.7}$ (eq.~[\ref{eq:cooling}]) is thermally unstable.  Radiative cooling can be stabilized, however, by the other source of (stable) wind cooling, given by the last term $\propto T$ in the entropy equation (eq.~[\ref{eq:dsdt}]), which at the stagnation radius equals $q v_{\rm w}^{2}/2$.

An important ratio controlling the importance of thermal stability is thus
\begin{align}
\left.\frac{\dot{q}_{\rm cool}}{q v_{\rm w}^{2}/2}\right|_{\rm r{s}} \simeq
  \begin{cases}
   0.02\eta M_{\bullet,8}^{0.92}v_{500}^{-7.8}\chi^{-7.8}  &, \text{core}\\
   0.08\eta M_{\bullet,8}^{0.5}v_{500}^{-5.4}\chi^{-5.4}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhostarrs}) and (\ref{eq:rhors}).  We thus see that we require $v_{\rm w} \gtrsim $ 300 km s$^{-1}$ for stability, with a weak sensitivity on $\eta$ and $M_{\bullet}$.

@@

However, \citet{McCourt+12} argue that even if cooling is exactly balanced by heating everywhere, there is still a local thermal instability that operates to turn hot gas into cold clouds.  However, this local instability only causes a significant local drop-out of mass from the hot medium (to low-T clouds), if the cooling timescale,
\be
t_{\rm cool} \equiv \frac{3n kT/2}{\dot{q}_{\rm cool}}
\ee 
is locally shorter than the approximately 10 times the free-fall timescale $t_{\rm ff}$.

@@

{\bf BDM: in fact, I think the above criteria for tcool/tff and qdotcool/qdotwind are identical to constants of order unit, since we are using gas density estimated from free-fall time... since the BH mass dependencies are different we need to recheck algebra}
